\newpage

## Area Allocation of Harvest

The apportionment of harvest for `r year` and `r year+1` was conducted using the REMA model using the same assumptions as in 2021:

  1. Because Amendment 41 prohibited trawling in the Eastern area east of 140째 W longitude, the ABC derived from REMA is split between W. Yakutat and E. Yakutat/SE using a weighted average of area-specific biomass ratios (see [Area Allocation of Harvests ](#apportionment) for a description of this approach).

  2. Because concerns regarding stock structure differences in the SEO vs other areas, the OFL for the SEO region remains separate.
 
The following table shows the recommended apportionment for 
The apportionment percentages are the same as in the last full assessment.  

```{r}
# apportionment table example
appo <- safe::apportion

flextable::flextable(appo[-1,]) %>% 
    flextable::add_header_row(values = paste0(appo[1,], "%"), top=FALSE) %>%
    flextable::border_remove() %>% 
    flextable::compose(i=1, j=1:2, part='header', 
                       value=flextable::as_paragraph('')) %>%
    flextable::compose(i=2, j=1:2, part='header', 
                        value=flextable::as_paragraph('Area Apportionment')) %>% 
    flextable::merge_h(i=1:2, part = "header") %>% 
    flextable::font(fontname = "Times New Roman", part = "all") %>%
    flextable::fontsize(size = 11, part = "all") %>% 
    flextable::align(j = 1, part = "all", align="left") %>% 
     flextable::compose(i = 1:2, j = 1, 
                       value = flextable::as_paragraph(as.character(year)), 
                       part = "body") %>% 
    flextable::compose(i = 3:4, j = 1, 
                       value = flextable::as_paragraph(as.character(year+1)), 
                       part = "body")  %>% 
    flextable::compose(i = 1:4, j = 2, 
                       value = flextable::as_paragraph(rep(c('ABC (t)', 'OFL (t)'), 2)), 
                       part = "body") %>% 
    # flextable::height(height = 0.2, part='body') %>% 
    # flextable::hrule(rule = "exact") %>% 
    flextable::hline_top(part='all') %>% 
    flextable::hline_bottom() 
  
```

Amendment 41 prohibited trawling in the Eastern area east of 140째 W longitude. 
The ratio of biomass still obtainable in the W. Yakutat area (between 147째 W and 140째 W) is 0.75. 
This results in the following apportionment to the W. Yakutat area:

```{r}


apportion2 <- safe::apportion2
flextable::flextable(apportion2) %>% 
  flextable::border_remove() %>% 
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  flextable::fontsize(size = 11, part = "all") %>%
  flextable::compose(i = 1, j = 1, 
                       value = flextable::as_paragraph(as.character(year)), 
                       part = "body") %>% 
  flextable::compose(i = 2, j = 1, 
                       value = flextable::as_paragraph(as.character(year+1)), 
                       part = "body") %>% 
  flextable::compose(i = 1:2, j = 2, 
                       value = flextable::as_paragraph(rep('ABC (t)', 2)), 
                       part = "body") %>% 
  flextable::compose(j=1:2, part='header', 
                      value=flextable::as_paragraph('')) %>% 
  flextable::compose(j=3, part='header', 
                      value=flextable::as_paragraph('W. Yakutat')) %>% 
    flextable::compose(j=4, part='header', 
                      value=flextable::as_paragraph('E. Yakutat/Southeast')) %>% 
  flextable::width(j = 3, width = 1) %>% 
  flextable::width(j = 4, width = 1.5) %>% 
          flextable::hline_top(part='all') %>% 
    flextable::hline_bottom() %>% 
  flextable::align(j = 1, part = "body", align="left")
  
```

\newpage
