\newpage

## Area Allocation of Harvest

The apportionment of harvest for `r year+1` and `r year+2` was conducted using the REMA model using the same assumptions as in 2021:

1. Because Amendment 41 prohibits trawling in the Eastern area east of 140째 W longitude, the ABC allocation derived from REMA for the Eastern Gulf is split between W. Yakutat and E. Yakutat/Southeast Outside ('Southeast' or 'SEO') using a weighted average of area-specific biomass ratios obtained from AKFIN (see [Area Allocation of Harvests ](#apportionment) for a description of this approach).

2. The OFL for the SEO region remains separate because of concerns regarding stock structure differences in the SEO vs other areas. Per the 2012 Groundfish Plan team, OFLs are combined for the Western, Central, and West Yakutat areas (W/C/WYK) because the original rationale of an overfished stock no longer applied. 

```{r}
## load three relevant tables to env(abc_apport, abc_apport_wyax, ofl_apport)
lapply(list.files(here::here(year, 'mgmt',model,'processed'),pattern = '_apport',full.names = TRUE), 
       FUN = function(x){
         y <- read.csv(x) 
         assign(x=gsub('.csv','',basename(x)), value  =y, envir = .GlobalEnv);rm(y);rm(x)
       }
)

## manually bind correct values
## ofl_apport is missing year 1, back it out using percentages
ofl_to_eyak <- round((1-abc_apport_wyak$wyak_p)*ofl_apport$ofl*abc_apport[1,'egoa'])
ofl_apport[2,3:4] <- c(ofl_apport$ofl[2]-ofl_to_eyak[2], ofl_to_eyak[2])

appo0 <- data.frame(year = as.character(c(year+1,year+2)),
                    item = c(rep('ABC (t)',2),rep('OFL (t)',2)),
                    WGOA = c(abc_apport[2:3,'wgoa'],rep(0,2)),
                    CGOA = c(abc_apport[2:3,'cgoa'],rep(0,2)),
                    EGOA_wyak = unlist(c(abc_apport_wyak['wyak'],ofl_apport['ofl_wcgoa_wyak'])),
                    EGOA_seo = unlist(c(abc_apport_wyak[1:2,'eyak_se'],ofl_apport['ofl_eyak_se'])),
                    total = unlist(c(abc_apport[2:3,'goa'],ofl_apport['ofl']))) 

appo0[3:4,'WGOA'] <- appo0[3:4,'EGOA_wyak'] ## move to LHS for formatting

flextable::flextable(appo0) %>% 
  ## format OFL stuff
  flextable::merge_at( i = 3, j = 3:5, part = 'body') %>%
  flextable::merge_at( i = 4, j = 3:5, part = 'body') %>%
  flextable::align(i = 1, align = 'center', part = 'header')  %>%
  flextable::align(i = 3:4, j = 3, align = 'center', part = 'body') %>% 
  
  flextable::add_header_row(values = paste0(c('REMA-derived apportionment (%)', 
                                              100*abc_apport[1,'wgoa'],
                                              100*abc_apport[1,'cgoa'],
                                              100*abc_apport[1,'egoa'],
                                              100*abc_apport[1,'goa'])), 
                            colwidths = c(2,1,1,2,1)) %>%
  flextable::set_header_labels(values = c('Year','','Western','Central','W. Yakutat','E. Yakutat/SE','Total')) %>%
  flextable::colformat_double(big.mark = ',', j = 3:7, digits = 0) %>%
  flextable::vline(j = c(2,3,4,6), part = 'body')%>%
  flextable::hline(i = c(2), part = 'body') %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  flextable::fontsize(size = 11, part = "all")  %>%
  flextable::bg(i = c(1,3), bg = 'grey88') %>% 
  flextable::hline_top(part='all') %>% 
  flextable::hline_bottom() 

```

Amendment 41 prohibited trawling in the Eastern area east of 140째 W longitude. 
The ratio of biomass still obtainable in the W. Yakutat area (between 147째 W and 140째 W) is 0.75. 
This results in the following apportionment to the W. Yakutat area:

```{r}


apportion2 <- safe::apportion2
flextable::flextable(apportion2) %>% 
  flextable::border_remove() %>% 
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  flextable::fontsize(size = 11, part = "all") %>%
  flextable::compose(i = 1, j = 1, 
                     value = flextable::as_paragraph(as.character(year)), 
                     part = "body") %>% 
  flextable::compose(i = 2, j = 1, 
                     value = flextable::as_paragraph(as.character(year+1)), 
                     part = "body") %>% 
  flextable::compose(i = 1:2, j = 2, 
                     value = flextable::as_paragraph(rep('ABC (t)', 2)), 
                     part = "body") %>% 
  flextable::compose(j=1:2, part='header', 
                     value=flextable::as_paragraph('')) %>% 
  flextable::compose(j=3, part='header', 
                     value=flextable::as_paragraph('W. Yakutat')) %>% 
  flextable::compose(j=4, part='header', 
                     value=flextable::as_paragraph('E. Yakutat/Southeast')) %>% 
  flextable::width(j = 3, width = 1) %>% 
  flextable::width(j = 4, width = 1.5) %>% 
  flextable::hline_top(part='all') %>% 
  flextable::hline_bottom() %>% 
  flextable::align(j = 1, part = "body", align="left")

```

\newpage
