## Harvest recommendations
 

### Amendment 56 Reference Points
Amendment 56 to the GOA Groundfish Fishery Management Plan defines the “overfishing level”
(OFL), the fishing mortality rate used to set OFL (FOFL), the maximum permissible ABC, and the fishing mortality rate used to set the maximum permissible ABC. The fishing mortality rate used to set ABC (FABC) may be less than this maximum permissible level, but not greater. Because reliable estimates of reference points related to maximum sustainable yield (MSY) are currently not available but reliable estimates of reference points related to spawning per recruit are available, POP in the GOA are managed under Tier 3 of Amendment 56. Tier 3 uses the following reference points: $B_{40}$%, equal to 40% of the equilibrium spawning biomass that would be obtained in the absence of fishing; $F_{35}$%,,equal to the fishing mortality rate that reduces the equilibrium level of spawning per recruit to 35% of the level that would be obtained in the absence of fishing; and $F_{40}$%, equal to the fishing mortality rate that reduces the equilibrium level of spawning per recruit to 40% of the level that would be obtained in the absence of fishing.

Estimation of the $B_{40}$% reference point requires an assumption regarding the equilibrium level of recruitment. In this assessment, it is assumed that the equilibrium level of recruitment is equal to the average of age-2 recruitments between 1979 and 2019 (i.e., the 1977 – 2017 year classes). Because of uncertainty in very recent recruitment estimates, we lag 2 years behind model estimates in our projection. Other useful biomass reference points which can be calculated using this assumption are $B_{100}$% and $B_{35}$%, defined analogously to $B_{40}$%. The 2022 estimates of these reference points are: 


 
|$B_{100}$%  |	331,917|
|$B_{40}$%	 | 132,767|
|$B_{35}$%	 | 116,171|
|$F_{40}$% | 0.10|
|$F_{35}$%	 | 0.12|




### Specification of OFL and Maximum Permissible ABC

Female spawning biomass for 2022 is estimated at 216,635 t. This is above the B40% value of 132,767 t. Under Amendment 56, Tier 3, the maximum permissible fishing mortality for ABC is F40% and fishing mortality for OFL is F35%. Applying these fishing mortality rates for 2022, yields the following ABC and OFL:

**TABLE HERE**


A standard set of projections is required for each stock managed under Tiers 1, 2, or 3 of Amendment 56.
This set of projections encompasses seven harvest scenarios designed to satisfy the requirements of Amendment 56, the National Environmental Policy Act, and the Magnuson-Stevens Fishery Conservation and Management Act (MSFCMA).

For each scenario, the projections begin with the vector of `r year` numbers at age as estimated in the assessment. 
This vector is then projected forward to the beginning of `r year + 1` using the schedules of natural mortality and selectivity described in the assessment and the best available estimate of total (year-end) catch for `r year`. 
In each subsequent year, the fishing mortality rate is prescribed on the basis of the spawning biomass in that year and the respective harvest scenario. 
In each year, recruitment is drawnfrom an inverse Gaussian distribution whose parameters consist of maximum likelihood estimates determined from recruitments estimated in the assessment. 
Spawning biomass is computed in each year based on the time of peak spawning and the maturity and weight schedules described in the assessment. 
Total catch after `r year` is assumed to equal the catch associated with the respective harvest scenario in all years. 
This projection scheme is run 1,000 times to obtain distributions of possible future stock sizes, fishing mortality rates, and catches.

Five of the seven standard scenarios will be used in an Environmental Assessment prepared in conjunction with the final SAFE. 
These five scenarios, which are designed to provide a range of harvest alternatives that are likely to bracket the final TAC for 2019, are as follow (“$max F_{ABC}$” refers to the maximum permissible value of $F_{ABC}$ under Amendment 56):

 - Scenario 1: In all future years, *F* is set equal to $max F_{ABC}$. (Rationale: Historically, TAC has been constrained by ABC, so this scenario provides a likely upper limit on future TACs.)
 
 - Scenario 2: In `r year` and `r year + 1`, *F* is set equal to a constant fraction of $max F_{ABC}$, where this fraction is equal to the ratio of the realized catches in `r year-3`-`r year-1` to the ABC recommended in the assessment for each of those years. 
For the remainder of the future years, maximum permissible ABC is used. (Rationale: In many fisheries the ABC is routinely not fully utilized, so assuming an average ratio catch to ABC will yield more realistic projections.)

 - Scenario 3: In all future years, *F* is set equal to 50% of $max F_{ABC}$. (Rationale: This scenario provides a likely lower bound on FABC that still allows future harvest rates to be adjusted downward when stocks fall below reference levels.)

 - Scenario 4: In all future years, *F* is set equal to the 2013-2017 average *F*. (Rationale: For some stocks, TAC can be well below ABC, and recent average *F* may provide a better indicator of $F_{TAC}$ than $F_{ABC}$.)

 - Scenario 5: In all future years, *F* is set equal to zero. (Rationale: In extreme cases, TAC may be set at a level close to zero.)

Two other scenarios are needed to satisfy the MSFCMA’s requirement to determine whether a stock is currently in an overfished condition or is approaching an overfished condition. 
These two scenarios are as follows (for Tier 3 stocks, the MSY level is defined as $B_{35\%}$):

 - Scenario 6: In all future years, *F* is set equal to $F_{OFL}$. (Rationale: This scenario determines whether a stock is overfished. If the stock is expected to be 1) above its MSY level in 2018 or 2) above ½ of its MSY level in 2018 and above its MSY level in 2028 under this scenario, then the stock is not overfished.)

 - Scenario 7: In `r year` and `r year + 1`, *F* is set equal to max $F_{ABC}$, and in all subsequent years F is set equal to FOFL. (Rationale: This scenario determines whether a stock is approaching an overfished condition. If the stock is 1) above its MSY level in 2020 or 2) above 1/2 of its MSY level in 2020 and expected to be above its MSY level in 2030 under this scenario, then the stock is not approaching an overfished condition.)
 
Spawning biomass, fishing mortality, and yield are tabulated for the seven standard projection scenarios (Table 10.16). The difference for this assessment for projections is in Scenario 2 (Author’s *F*); we use pre-specified catches to increase accuracy of short-term projections in fisheries where the catch is usually less than the ABC. This was suggested to help management with setting preliminary ABCs and OFLs for two-year ahead specifications.

In addition to the seven standard harvest scenarios, Amendments 48/48 to the BSAI and GOA Groundfish Fishery Management Plans require projections of the likely OFL two years into the future. 
While Scenario 6 gives the best estimate of OFL for `r year`, it does not provide the best estimate of OFL for `r year + 1`, because the mean `r year` catch under Scenario 6 is predicated on the `r year` catch being equal to the `r year` OFL, whereas the actual `r year` catch will likely be less than the `r year` OFL. 
The executive summary contains the appropriate one- and two-year ahead projections for both ABC and OFL.


During the 2006 CIE review, it was suggested that projections should account for uncertainty in the entire assessment, not just recruitment from the endpoint of the assessment. We continue to present an alternative projection scenario using the uncertainty of the full assessment model, harvesting at the same estimated yield ratio as Scenario 2, except for all years instead of the next two. This projection propagates uncertainty throughout the entire assessment procedure based on MCMC. The projection shows wide credibility intervals on future spawning biomass (Figure 9-17). The $B_{35\%}$ and $B_{40\%}$ reference points and future recruitment is based on the 1979-2019 age-2 recruitment, and this projection predicts that the median spawning biomass will eventually tend toward these reference points while at harvesting at  $F_{40\%}$.


## Risk Table and ABC recommendation


The SSC in its December 2018 minutes recommended that all assessment authors use the risk table when determining whether to recommend an ABC lower than the maximum permissible. The following template is used to complete the risk table:

```{r}
library(flextable)
library(tidyverse)

data.frame(level = c("Level 1: Normal",
                     "Level 2: Substantially increased concerns",
                     "Level 3: Major Concern",
                     "Level 4: Extreme concern"),
           assess = c("Typical to moderately increased uncertainty/minor unresolved issues in assessment.", 
                      "Substantially increased assessment uncertainty/ unresolved issues.",
                      "Major problems with the stock assessment; very poor fits to data; high level of uncertainty; strong retrospective bias.",
                      "Severe problems with the stock assessment; severe retrospective bias. Assessment considered unreliable."),
           pop = c("Stock trends are typical for the stock; recent recruitment is within normal range.",
                   "Stock trends are unusual; abundance increasing or decreasing faster than has been seen recently, or recruitment pattern is atypical.",
                   "Stock trends are highly unusual; very rapid changes in stock abundance, or highly atypical recruitment patterns.",
                   "Stock trends are unprecedented; More rapid changes in stock abundance than have ever been seen previously, or a very long stretch of poor recruitment compared to previous patterns."),
           enviro = c("No apparent environmental/ecosystem concerns",
                      "Some indicators showing adverse signals relevant to the stock but the pattern is not consistent across all indicators.",
                      "Multiple indicators showing consistent adverse signals a) across the same trophic level as the stock, and/or b) up or down trophic levels (i.e., predators and prey of the stock)",
                      "Extreme anomalies in multiple ecosystem indicators that are highly likely to impact the stock; Potential for cascading effects on other ecosystem components"),
           fish = c("No apparent fishery/resource-use performance and/or behavior concerns",
                    "Some indicators showing adverse signals but the pattern is not consistent across all indicators",
                    "Multiple indicators showing consistent adverse signals a) across different sectors, and/or b) different gear types",
                    "Extreme anomalies in multiple performance  indicators that are highly likely to impact the stock")) %>% 
  flextable() %>% 
    font(fontname = "Times", part = "all") %>% 
    width(j = ~ assess, width = 1.5) %>% 
    width(j = ~ pop, width = 2) %>% 
    width(j = ~ enviro, width = 1.5) %>% 
    width(j = ~ fish, width = 1.5) %>% 
    set_header_labels(level = "",
                    assess = "Assessment-related considerations",
                    pop = "Population dynamics considerations",
                    enviro = "Environmental/ecosystem considerations",
                    fish = "Fishery Performance") %>% 
  font(fontname = "Times", part = "all") %>% 
  fontsize(size = 10) %>% 
  theme_vanilla() %>% 
  italic(part = "header") 
```  
The table is applied by evaluating the severity of four types of considerations that could be used to support a scientific recommendation to reduce the ABC from the maximum permissible. 
These considerations are stock assessment considerations, population dynamics considerations, environmental/ecosystem considerations, and fishery performance. Examples of the types of concerns that might be relevant include the following: 

  1. “Assessment considerations—data-inputs: biased ages, skipped surveys, lack of fishery-independent trend data; model fits: poor fits to fits to fishery or survey data, inability to simultaneously fit multiple data inputs; model performance: poor model convergence, multiple minima in the likelihood surface, parameters hitting bounds; estimation uncertainty: poorly-estimated but influential year classes; retrospective bias in biomass estimates.
  2. “Population dynamics considerations—decreasing biomass trend, poor recent recruitment, inability of the stock to rebuild, abrupt increase or decrease in stock abundance.
  3. “Environmental/ecosystem considerations—adverse trends in environmental/ecosystem indicators, ecosystem model results, decreases in ecosystem productivity, decreases in prey abundance or availability, increases or increases in predator abundance or productivity.
  4. “Fishery performance—fishery CPUE is showing a contrasting pattern from the stock biomass trend, unusual spatial pattern of fishing, changes in the percent of TAC taken, changes in the duration of fishery openings.”



#### Assessment considerations

In recent assessments the GOA POP assessment model has resulted in a negative retrospective pattern, which is interpreted as the model continually increasing spawning biomass as new data are added (-0.16 in the current assessment, Figure 9-17). While the assessment fits to composition data from the survey (age) and fishery (age and length) are generally adequate (Figures 9-2, 9-3, and 9-6), the retrospective pattern is driven by increases in the trawl survey biomass estimates since 2013. The assessment model has underestimated each survey biomass estimate since 2013, resulting in five consecutive years of negative residuals (Figure 9-4). This consistent retrospective pattern across recent assessments could indicate misspecification in the model, and thus, we continue to rank the assessment considerations as level 2, a substantially increased concern.


#### Population dynamics considerations

As discussed in the **Assessment considerations** section above, the recent increase in POP biomass since 2011 is an unusual increase that has not been seen in the time series of biomass prior. In order to fit these large bottom trawl survey biomass estimates the assessment model has indicated several above average recruitment events in recent years (Figures 9-13 and 9-14), most notably in the mid-1980s, mid- and late-1990s, and since 2000. However, even with these above average recruitments the model is still not able to fit the increase bottom trawl survey biomass satisfactorily. In comparison to many stocks in the North Pacific, this increase in biomass coinciding with warmer temperatures is atypical (with the exception of sablefish). This stock trend is unusual because both the stock trend and recruitment estimates have been increasing faster than seen recently, and as such, we continue to rate the population-dynamics concern as level 2, a substantially increased concern.

#### Environmental/Ecosystem considerations

We scored this category as level 1 (normal concern) for Pacific ocean perch (POP) given moderate thermal conditions for adults and moderate to below average thermal conditions for larvae, mixed trends for zooplankton abundance, and potential but unknown levels of competition with juvenile sablefish and pink salmon.
POP are benthic, continental slope (150-300 m depths) dwellers as adults, with a pelagic then inshore benthic juvenile stage (age 1 to 3) in the Gulf of Alaska (GOA) (Carlson and Haight 1976, Love et al. 2002, Rooper and Bolt 2005, Rooper et al. 2007, NPFMC 2010). Spawning occurs during winter and early spring , larvae settle to the benthos within 3-6 months (Love et al. 2002). It is reasonable to expect that the 2021 and predicted 2022 average deeper ocean temperatures will provide good spawning habitat. Average to cooler surface temperatures will contribute to average to below average pelagic conditions for larval rockfish, as warm spring surface waters are conducive to larval survival and positive rockfish recruitment (Moss 2016, Morgan 2019). Ocean temperatures at the surface and at depth on the shelf were around the long-term average in 2021 (AFSC Bottom Trawl Survey, Laman 2021; AFSC EcoFOCI survey, Rogers et al 2021; Seward Line Survey,  Danielson 2021), although western GOA started the year with warmer surface waters (satellite data; Watson 2021) and there was slightly above average warmth (5.2 °C) at 200m depth along the outer edge of the shelf during the summer (AFSC Longline Survey; Siwicke 2021). This is within the range of preferred ocean temperatures for adults (4.0-6.5 °C, Major and Shippen 1970). Numerous temperature time series showed signs of cooling from previous surveys (returning to average from recent marine heatwave years 2014-2016, 2019) at the surface and at depth, and 2022 surface temperatures are predicted to continue cooling, in alignment with La Niña conditions and a negative Pacific Decadal Oscillation.
Planktivorous foraging conditions were moderate and regionally variable across the GOA in 2021. The primary prey of the adult Pacific Ocean Perch include calanoid copepods, euphausiids, myctophids, and miscellaneous prey in the GOA (Byerly 2001, Yang and Nelson 2000, Yang 2003). While zooplankton trends were variable, an important observation shows continued decline in POP body condition (i.e. lower weights at length) since 2015 (Bottom Trawl Survey, O’Leary, 2021). Declines are especially pronounced in SEAK and Yakutat regions since the previous survey in 2019. The timing of this declining trend matches the time frame of increasing POP population since the 2014-2016 marine heatwave. Regionally, zooplankton trends varied from the western to eastern GOA. The western GOA had lower spring biomass of large copepods and approximately average biomass of smaller copepods was around Kodiak, characteristics of previous warm, less productive years (e.g., 2019). Planktivorous seabird reproductive success, an indicator of zooplankton availability and nutritional quality, was below average just north of Kodiak (E. Amatuli Island; Drummond and Renner 2021). Around the eastern edge of WGOA (Seward Line, Middleton Island) the biomass of large copepods was average to above-average (Seward Line Survey, Hopcroft 2021) and planktivorous seabirds had better reproductive success (Middleton Island, Hatch et al. 2021), indicating improved forage conditions. The eastern GOA inside waters of Icy Strait, northern southeast Alaska, had higher than average large copepods and euphausiids (AFSC SECM Survey, Icy Strait, Fergusson 2021), however planktivorous seabirds had mixed reproductive success. Potential competitors are large year classes of juvenile sablefish (2016, 2018) and a pink salmon which are returning in very high numbers this year (Murphy et al. 2021, Shaul et al. 2021). 
Predators of juvenile POP include Pacific halibut, arrowtooth flounder, seabirds, rockfish, salmon, and lingcod (Moss 2016, Hulson et al. 2020). Predators of adults include Pacific halibut, sablefish, and sperm whales (Moss 2016, Hulson et al. 2020). Halibut and arrowtooth flounder populations remain low relative to previous levels, and, in general, there is no cause to suspect increased predation pressure on larval or adult demersal shelf rockfish. 


#### Fishery performance

In general, fishery CPUE shows consistent patterns in abundance similar to the bottom trawl survey and there have been no recent changes to spatial distribution of catch, percent of TAC taken, or fishing duration. Overall, there are no indications of adverse signals or concerns about the fishery in terms of resource-use, performance, or behavior and thus we scored the fishery-performance concern as level 1, no apparent concern. We will continue to monitor the fishery performance as it pertains to the COVID-19 pandemic.

#### Summary and ABC recommendation

The following is a summary of the risk table:
 



```{r}
l1 = "Level 1: No increased concerns"
l2 = "Level 2: Substantially increased concerns"
l3 = "Level 3: Major concern"
l4 = "Level 4: Extreme concern"


data.frame(assess = l2,
           pop = l2,
           enviro = l1,
           fish = l1) %>% 
  flextable() %>% 
    width(j = ~ assess, width = 1.5) %>% 
    width(j = ~ pop, width = 1.5) %>% 
    width(j = ~ enviro, width = 1.5) %>% 
    width(j = ~ fish, width = 1.5) %>% 
    set_header_labels(level = "",
                    assess = "Assessment-related considerations",
                    pop = "Population dynamics considerations",
                    enviro = "Environmental/ecosystem considerations",
                    fish = "Fishery Performance") %>% 
  fontsize(size = 10) %>% 
  font(fontname = "Times", part = "all") %>% 
  italic(part = "header") 
```

Bottom trawl survey estimates of POP biomass in the GOA indicate an unprecedented increase in abundance, which has not been properly explained by the population dynamics defined in the current assessment model. Even though we rate the assessment-related and population dynamics considerations at a level 2, we do not, however, recommend a reduction in ABC as the retrospective pattern in this assessment continues to indicate increasing population abundance.



### Area Allocation of Harvests

Apportionment of ABC and OFL among regulatory areas has been based on the random effects model developed by the survey averaging working group. The random effects model was fit to the survey biomass estimates (with associated variance) for the Western, Central, and Eastern GOA. The random effects model estimates a process error parameter (constraining the variability of the modeled estimates among years) and random effects parameters in each year modeled. The fit of the random effects model to survey biomass in each area is shown in Figure 9-18.
In general the random effects model fits the area-specific survey biomass reasonably well. The random effects model estimates increases in biomass in the Western and Central Gulf in 2021 compared to 2019, and decrease in the Eastern Gulf. Using the random effects model estimates of survey biomass for the apportionment results in 6.8% for the Western area (up from 4.6% in 2019), 80.5% for the Central area (up from 75.8% in 2019), and 12.7% for the Eastern area (down from 19.6% in 2019).
Using the results of the random effects model results in recommended ABC’s of 2,602 t for the Western area, 30,806 t for the Central area, and 4,860 t for the Eastern area.
Amendment 41 prohibited trawling in the Eastern area east of 140° W longitude. In the past, the Plan Team has calculated an apportionment for the West Yakutat area that is still open to trawling (between 147oW and 140oW). We calculated this apportionment using the ratio of estimated biomass in the closed area and open area. This calculation was based on the team’s previous recommendation that we use the weighted average of the upper 95% confidence interval for the W. Yakutat. We computed this interval this year using the weighted average of the ratio for 2017, 2019, and 2021. We calculated the approximate upper 95% confidence interval using the variance of a weighted mean for the 2017-2021 weighed mean ratio. This resulted in a ratio of 0.29, up from 0.24 in 2019. This results in an ABC apportionment of 1,409 t to the W. Yakutat area which would leave 3,451 t unharvested in the Southeast/Outside area.
Based on the definitions for overfishing in Amendment 44 in tier 3a (i.e., FOFL = F35%=0.12), overfishing is set equal to 45,580 t for POP. The overfishing level is apportioned by area for POP and historically used the apportionment described above for setting area specific OFLs. However, in 2012, area OFLs were combined for the Western, Central, and West Yakutat (W/C/WYK) areas, while East Yakutat/Southeast (SEO) was separated to allow for concerns over stock structure. This results in overfishing levels for W/C/WYK area of 41,470 t and 4,110 t in the SEO area. 


### Status Determination

Under the MSFCMA, the Secretary of Commerce is required to report on the status of each U.S. fishery with respect to overfishing. 
This report involves the answers to three questions: 1) Is the stock being subjected to overfishing? 2) Is the stock currently overfished? 3) Is the stock approaching an overfished condition?

*Is the stock being subjected to overfishing?* The official catch estimate for the most recent complete year (`r year-1`) is *correct this later* r catch %>% filter(Year==year-1) %>% pull(Catch) %>% format(., big.mark = ",")` t. 
This is less than the `r year-1` OFL of 5,402 t. 
Therefore, the stock is not being subjected to overfishing.

Harvest Scenarios #6 and #7 are intended to permit determination of the status of a stock with respect to its minimum stock size threshold (MSST). 
Any stock that is below its MSST is defined to be overfished. 
Any stock that is expected to fall below its MSST in the next two years is defined to be approaching an overfished condition. 
Harvest Scenarios #6 and #7 are used in these determinations as follows:

*Is the stock currently overfished?* This depends on the stock’s estimated spawning biomass in `r year`:

 - a. If spawning biomass for `r year` is estimated to be below ½ $B_{35\%}$, the stock is below its MSST.
 - b. If spawning biomass for `r year` is estimated to be above $B_{35\%}$ the stock is above its MSST.
 - c. If spawning biomass for `r year` is estimated to be above ½ $B_{35\%}$ but below $B_{35\%}$, the stock’s status relative to MSST is determined by referring to harvest Scenario #6 (Table 10.16). 
If the mean spawning biomass for 2028 is below $B_{35\%}$, the stock is below its MSST. 
Otherwise, the stock is above its MSST.

*Is the stock approaching an overfished condition?* This is determined by referring to harvest Scenario #7:

 - a. If the mean spawning biomass for `r year+2` is below 1/2 $B_{35\%}$, the stock is approaching an overfished condition.
 - b. If the mean spawning biomass for `r year+2` is above $B_{35\%}$, the stock is not approaching an overfished condition.
 - c. If the mean spawning biomass for `r year+2` is above 1/2 $B_{35\%}$ but below $B_{35\%}$, the determination depends on the mean spawning biomass for `r year+12` 
If the mean spawning biomass for `r year+12` is below $B_{35\%}$, the stock is approaching an overfished condition. 
Otherwise, the stock is not approaching an overfished condition.
Based on the above criteria and Table 10.16, the stock is not overfished and is not approaching an overfished condition.

The fishing mortality that would have produced a catch for last year equal to last year’s OFL is `r best_f`.
