# Executive Summary

Gulf of Alaska `r species` (*Sebastes alutus*) have historically been assessed on a biennial stock assessment schedule to coincide with the availability of new trawl survey data (odd years). 
In 2017, the Alaska Fisheries Science Center (AFSC) participated in a stock assessment prioritization process. 
It was recommended that Gulf of Alaska (GOA) `r species` remain on a biennial stock assessment schedule with a full stock assessment produced in even years and a partial stock assessment produced in odd years. 
For a partial assessment, the projection model is updated with new catch information and results are used to recommend harvest levels for the next two years. 
This incorporates the most current catch information without re-estimating model parameters and biological reference points.

Gulf of Alaska `r species` are classified as a Tier 3 stock and are assessed using a statistical age-structured model.
This assessment consists of a population model, which uses survey and fishery data to generate a historical time series of population estimates, and a projection model, which uses results from the population model to predict future population estimates and recommended harvest levels. 
The data used in this assessment includes total catch biomass, fishery age and size compositions, trawl survey abundance estimates, and trawl survey age compositions. 

## Summary of Changes in Assessment Inputs

*Changes in the input data:* There were no changes made to the assessment model inputs as this is an off-cycle year. 
New data added to the projection model included updated catch data from `r year-1` (`r c0` t) and new estimated catches for `r year`-`r year + 2`. 
Catch data were queried on `r data_pull`. 
The `r year` catch was estimated by increasing the observed catch by an expansion factor of `r round(yld$ratio, 3)`, which accounts for the average fraction of catch taken after `r data_pull2` in the last three complete years (`r year - 3`-`r year-1`). 
This expansion factor increased from last year’s expansion factor of 1.036 and resulted in an estimated catch for `r year` of `r c1` t. 
To estimate future catches, we updated the yield ratio to `r round(yld$yld, 2)`, which was the average ratio of catch to ABC for the last three complete catch years. 
This yield ratio was multiplied by the projected ABCs from the updated projection model to generate catches of `r c2` t in `r year + 1` and `r c3` t in `r year + 2`.


*Changes in assessment methodology:* There were no changes from the `r year-1` assessment [@Williams2022] since this is an off-cycle year.


## Summary of Results

*ABC recommendation*  
The projected total biomass for `r year+1` is `r tot` t. 
The recommended ABC for `r year+1` is `r abc` t, the maximum allowable ABC under Tier `r tier`. 
This ABC is a `r perc`% decrease compared to the `r year` ABC of `r abc0` and a 1% increase from the projected 2024 ABC from the last full  assessment.
The 2024 GOA-wide OFL for `r species` is `r ofl` t.
The stock is not being subject to overfishing, is not currently overfished, nor is it approaching a condition of being overfished. 

Reference values for GOA `r species` are summarized in the following table, with the recommended ABC and OFL values in bold.

```{r, child= "01a-main_table.Rmd", eval = T}

```


```{r, child= "01b-apportionment.Rmd", eval = T}

```
